The mass distribution of companions to low-mass white dwarfs

### 2014-10-22 documents/letter/ms.tex
Updated .tex

\title{The mass distribution of companions to extremely low-mass white dwarfs}
\author{Jeff Andrews\altaffilmark{\colum}, Adrian M. Price-Whelan\altaffilmark{\colum}, Marcel Ag\"ueros\altaffilmark{\colum}}

@@

\begin{abstract}
Measuring the masses of companions to single-line spectroscopic binary stars
is (in general) not possible because of the unknown orbital plane inclination angles. 
Even when the mass of the visible star can be measured, only a lower limit can be placed
on the mass of the unseen companion. However, since these inclination angles should be isotropically
distributed, for a large enough, unbiased sample, the 
companion mass distribution can be deconvolved from the distributions of observables.
In this work, we construct a hierarchical probabilistic model to infer properties of unseen 
companion stars to low mass white dwarfs (LMWD) given orbital period and projected radial velocity measurements of the primary. We use a mixture of 
two Gaussians to model the true companion mass distribution to represent 
populations of white dwarfs (WD) and neutron stars (NS). 
Our model successfully recovers the initial parameters of several mock data sets.
We then apply our model to 55 of the 61 white dwarfs in 
the extremely low mass (ELM) WD survey. We find maximum a posteriori (MAP) parameters for
the WD companion population of $\mu_{\rm WD} = 0.71~\Msun$ and $\sigma_{\rm WD} = 0.26~\Msun$ 
with a NS fraction consistent with zero, $f_{\rm NS} = 0\%$. However, the marginal posterior 
distribution over NS fraction has a significant tail up to $f_{\rm NS} \approx 16\%$.
We make samples from the posterior distribution publicly available so that future observational
efforts may compute the NS probability for newly discovered low mass WDs.
\end{abstract}

@@

Except in special cases of extreme metallicity \citep{kilic07}, the Galaxy is not old enough to produce WDs with $M\lesssim0.45~ \Msun$ through single star evolution. These objects are formed through dynamical interactions with another star, and are therefore expected to be found in binary systems. \citep{han98,nelemans00,nelemans01,vdSluys06,woods12}. In their seminal work, \citet{marsh95} discovered companions around five LMWDs in their sample of seven objects, confirming these theoretical expectations. The ELM WD Survey identifies LMWDs in SDSS for spectroscopic follow-up \citep{ELMI}. After their initial investigation found 11 of 12 LMWDs had close binary companions, further spectroscopic follow-up has now brought the sample of ELM WDs to 61 systems, 55 of which have measured radial velocity amplitudes and orbital periods \citep{ELMII, ELMIII, ELMIV, ELMV}.

While the orbital periods and radial velocities indicated the companions were most likely WDs, due to the unmeasured inclination angle, these LMWDs could host NS companions \citep{vLeeuwen07}. However, all radio searches for pulsed emission, as well as searches for blackbody X-ray emission from a NS surface have been unsuccessful \citep{agueros09a,agueros09b,kilic13}. Yet, such systems must exist since LMWDs are observed as companions around millisecond pulsars \citep{vKerkwijk96,callanan98,bassa06,antoniadis12}. Finding even one system would be extremely interesting: the NS binaries with WD companions bright enough to detect spectroscopic radial velocity variations have precisely determined NS masses which provide useful insights into the binary evolution processes that form these systems \citep{benvenuto05,smedley14,jia14}

Selecting the most likely candidates for multiwavelength follow-up is crucial. For each WD in a single-line spectroscopic binary, spectroscopic observations provide the orbital period ($\period$), primary WD mass ($M_1$), and projected orbital velocity ($K=v \sin i$). Using Kepler's third law, we can write the relation between these quantities and the unknown inclination angle as:
\begin{equation}
	\frac{(M_2 \sin i)^3}{\left(M_1+M_2\right)^2} = \frac{\period}{2\pi G} K^3 = \mf \label{eq:massfunc}
\end{equation}
The righthand side of this equation is the mass function ($\mf$). $M_2$ is minimized for an edge-on orbit, with $i = 90\degree$. Because of the dependence on $i$, the nature of the companion cannot usually be determined based on $\mf$ alone. Figure~\ref{fig:Porb-M1} shows that the population of LMWD with pulsar companions (triangles) occupy the same region in $M_1 - \period$ space as those with WD companions (circles). Therefore, barring rare circumstances (such as an edge-on, eclipsing system), individual LMWDs with NS companions cannot be readily identified.

55 of the 61 systems have precise $M_1$, $\period$, and $K$ measured; we refer to this as the ELM sample. Since the population was selected based on characteristics determined by the primary mass alone, at least with regard to the inclination angle and companion mass, the population is unbiased. In this work, we develop a probabilistic model to infer parameters of an assumed form for the mass distribution of companions to ELM WDs. Our method is similar to that employed by \citet{ozel12} and \citet{kiziltan13} to describe the mass distribution of NSs in binaries. We focus on several specific astrophysical questions: 1.\ Can the population as a whole be modeled using a simple description of the companion masses? 2.\ How does the companion mass distribution compare to predictions from population synthesis simulations?  3.\ What is the rate of NS-LMWD binaries implied by our model? 4.\ What are the resulting distributions of NS probabilities for individual systems in the ELM sample?

@@

where the product is over each of the $j$ systems.  This marginal likelihood involves integrals over the latent variables $i$ and $M_2$,
\begin{align}
    p(\mf \given \bs{\theta}) &= \int_0^{\pi/2} di \int_{-\infty}^\infty dM_2  \nonumber \\
      & \qquad {} \times p(\mf \given M_1, M_2, i)~p(M_2 \given \bs{\theta})~p(i).
\end{align}
We neglect observational uncertainties in $\mf$ and $M_1$,\footnote{The fractional uncertainties in these quantities are small, $\sigma_x / x \sim 0.05-0.1$.} and assume the inclination angles are isotropically distributed
\begin{align}
	p(\mf \given M_1, M_2, i) &= \delta \left[\mf - f(M_1, M_2, i) \right]\\
	f(M_1, M_2, i) &= \frac{(M_2 \sin i)^3}{(M_1 + M_2)^2}\\
	p(i) &= \sin i.
\end{align}
For now, we keep this general by not yet specifying a parametric form for the companion mass distribution, $p(M_2 \given \bs{\theta})$. With the above assumptions, the marginal likelihood integral is
\begin{align}
%    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta}) \int_0^{\pi/2} di \nonumber \\
%    & \qquad {} \times \delta \left[\mf - f(M_1, M_2, i)\right]~\sin i\\
    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2 ~p(M_2 \given \bs{\theta}) \int_0^{\pi/2} di  \nonumber \\
    & \qquad {} \times ~\sin i ~ \delta \left[g(M_1,M_2,i) \right]\label{eq:delta}
\end{align}
where
\begin{equation}
	g(M_1,M_2,i) = \mf - \frac{M_2^3}{(M_1+M_2)^2}\sin^3 i.
\end{equation}
The inner integral (over $i$) has the form
\begin{equation}
    \int dx~F(x)~\delta \left[ G(x) \right] = \sum_j \frac{F(x^*_j)}{|G'(x^*_j)|}
\end{equation}
where the sum is over the roots, $x^*_j$, of the function $G(x)$. The root, $i^*$, and derivative of the argument of the delta function in \eqn\ref{eq:delta} are 
\begin{align}
%	\sin i^* &= \alpha\\
%	\alpha &= \frac{(\mf(M_1+M_2)^2)^{1/3}}{M_2}\\
	\sin i^* &= \frac{ \left[\mf(M_1+M_2)^2 \right]^{1/3}}{M_2}\\
%	\frac{\partial g}{\partial i} &= \frac{3M_2^3}{(M_1+M_2)^2}\sin^2 i \cos i\\
%	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \alpha^2 \sqrt{1 - \alpha^2}
	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \sin^2 i^* \sqrt{1 - \sin^2 i^*}
\end{align}
With this in mind, we may write the marginal likelihood as
\begin{align}
	p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\sin i^* \left(\frac{\partial g}{\partial i}\bigg\rvert_{i^*}\right)^{-1}\\
%	&= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\alpha\left[\frac{3M_2^3}{(M_1+M_2)^2} \alpha^2 \sqrt{1 - \alpha^2}\right]^{-1}\\
	&= \int_{M_{2,{\rm min}}}^\infty dM_2~p(M_2 \given \bs{\theta})~h(M_2, \mf, M_1) \label{eq:fullm2}
\end{align}

@@

We must now choose a functional form for the companion mass distribution,  $p(M_2\given \bs{\theta})$. For all experiments -- including that with the real data -- we use a two component Gaussian mixture model to fit the data. However, in tests below, we generate test data with different mixture component forms. We truncate the distributions using physically motivated bounds so that the white dwarf component is restricted to the range $M_2\in (0.2,\wdupper)~\Msun$ and the neutron star component is restricted to the range $M_2\in (1.3,2)~\Msun$,
\begin{align}
%	h(M_2, \mf, M_1) &= \frac{(M_1+M_2)^{4/3}}{3\mf^{1/3}M_2\sqrt{M_2^2 - \left[ \mf(M_1+M_2)^2 \right]^{2/3}}}\\
%	p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~\left[ (1-f_{\rm NS})p_{\rm WD} + f_{\rm NS}p_{\rm NS} \right]~h(M_2, \mf, M_1)\\
	p(M_2 \given \bs{\theta}) &= \left[ (1-f_{\rm NS})~p_{\rm WD} + f_{\rm NS}~p_{\rm NS} \right] 
\end{align}
where 
\begin{align}
	p_{\rm WD} &= \mathcal{N}(M_2 \given \mu_{\rm WD}, \sigma^2_{\rm WD});\,\left[ 0.2 < \frac{M_2}{\Msun} < \wdupper \right] \\
	p_{\rm NS} &= \mathcal{N}(M_2 \given \mu_{\rm NS}, \sigma^2_{\rm NS});\, \left[1.3 < \frac{M_2}{\Msun} < 2 \right].
\end{align}
and $\mathcal{N}$ is the (truncated, but properly normalized) normal distribution with mean $\mu$ and variance $\sigma^2$ and the distributions are limited to the ranges specified. In this work, we fix the mean and variance of the neutron star distribution to
\begin{align}
	\mu_{\rm NS} &= 1.4~\Msun\\
	\sigma^2_{\rm NS} &= (0.05~\Msun)^2
\end{align}
based on indications that NSs in certain binaries may be somewhat more massive than the canonical NS mass of 1.35 \Msun \citep{kiziltan13,smedley14}.

@@

Our companion mass model parameters are then $\bs{\theta} = (\mu_{\rm WD}, \sigma_{\rm WD}, f_{\rm NS})$. We choose standard priors on these parameters: for the mean of the white dwarf component, we use a uniform distribution from $0.2-1.4~\Msun$; we use a (scale-invariant) logarithmic prior for the variance of the white dwarf component over the range $0.05-2~\Msun$; finally, we use a uniform prior over the dimensionless neutron star fraction from $0-1$. The model parameters are summarized in Table~\ref{tbl:parameters}.

@@

We use an ensemble Markov Chain Monte Carlo algorithm \citep{goodman10} to draw samples from the posterior distribution, $p(\mu_{\rm WD}, \sigma_{\rm WD}, f_{\rm NS} \given \bs{m}_f, \bs{M}_1)$. The algorithm is implemented in the \texttt{Python} programming language \citep[{\tt emcee};][]{foremanmackey13} and uses an ensemble of individual ``walkers'' to naturally adapt to the geometry of the parameter-space being explored. We run the walkers for an initial, burn-in period of 500 steps starting from randomly drawn initial conditions (sampled from the priors in Table~\ref{tbl:parameters}). We then re-initialize the walkers from their positions at the end of this run, throw out the burn-in samples, and run again for 1000 steps. We compute the autocorrelation time for each parameter to confirm they have converged.

\section{Applying Our Model} \label{sec:tests}
To test the performance of this Gaussian mixture model, we first apply it to three separate sets of mock data (described in detail below), each with 100 ``observed'' systems. These data are generated by randomly drawing the companion mass, $M_2$, from a particular distribution, then using a randomly drawn inclination angle and primary mass, $M_1$, to compute the mass function, $\mf$. We draw primary masses from a uniform distribution, $M_1 \sim \mathcal{U}(0.2,0.4)~\Msun$. We apply the same Gaussian mixture model to all test datasets to infer the parameters of the white dwarf mixture component and neutron star fraction.

\begin{figure*}[h!]
\begin{center}
\includegraphics[width=0.95\textwidth]{many-panel.pdf}
\caption{Results from the three tests described in Section~\ref{sec:tests}. The left-most column shows histograms of $M_2$'s drawn from each of our test distributions. Overplotted as solid lines are the MAP models. The second column shows samples from the posterior distributions of $\mu_{\rm WD}$ and $\sigma_{\rm WD}$. The dashed lines in the top two panels show the true values that the sample systems were drawn from. Contours designate the 68\% and 95\% contours. Since the third model was not drawn from a Gaussian distribution, there is no true parameter set. The third panel shows posterior samples for $f_{\rm NS}$ and $\sigma_{\rm WD}$, again the dashed lines show the input values. Finally, the rightmost panels list the individual systems (ordered by increasing $\mf$) and show each of their corresponding NS probabilities. }
\label{fig:tests}
\end{center}
\end{figure*}

\subsection{Test 1: Single Gaussian (white dwarf)} \label{sec:exp1}

As a first test, we generate $M_2$'s by drawing from a single, truncated Gaussian with parameters given in Table~\ref{tbl:parameters}. This test sample contains no NSs. The leftmost panel in Figure~\ref{fig:ELM_post} shows that the maximum a posteriori (MAP) $M_2$ distribution (black line) qualitatively matches the input distribution (grey histogram). The second panel shows samples from the posterior distribution and contours that contain 68\% and 95\% of the samples. The input values lie cleanly within the inner contour in both projections of the posterior shown. The third panel shows that although the posterior $f_{\rm NS}$ distribution is consistent with 0\%, there is a tail up to $\sim$10\%. Nevertheless, our model recovers the input parameters for the white dwarf companion population. The fourth panel lists individual systems, ordered by $\mf$, and shows the distributions of $P_{\rm NS}$ for each. Few systems have $P_{\rm NS}$ values higher than $\sim$5\%.

\subsection{Test 2: Two Gaussians (white dwarf + neutron star)} \label{sec:exp2}
For our next test, we use the same Gaussian distribution to generate companion masses for the WD component as the previous example (Section~\ref{sec:exp1}), but we now add an additional NS component with $f_{\rm NS} = 10\%$ and parameters described in Section~\ref{sec:experiments}. The panels are the same as in Test 1 described above. Again, our model recovers the input values for $\mu_{\rm WD}$ and $\sigma_{\rm WD}$. Importantly, the third panel shows that it also recovers $f_{\rm NS}$, however the posterior again shows a tail toward higher probabilities. The fourth panel indicates that some systems have near unity $P_{\rm NS}$ values; our model correctly identifies these as true NS. However, this accounts for less than half of the systems with NSs in the simulated sample; the other half have low $\mf$'s due to low inclination angles.

\subsection{Test 3: Uniform (white dwarf) + Gaussian (neutron star)} \label{sec:exp3}
For a final test, we generate companion masses for the WD component by sampling from a uniform distribution over the range $(0.2,1.2)~\Msun$, while keeping $f_{\rm NS}=$10\%. The panels are the same as in Test 1 described above. The posterior distribution in the second panel shows that the systems have difficulty converging on any one solution for $\mu_{\rm WD}$ and $\sigma_{\rm WD}$. Interestingly, the third panel shows that despite having a non-Gaussian input distribution for $M_2$, our model still recovers $f_{\rm NS}$ approximately as accurately as in Test 2.

@@

The ELM WD sample is composed of 55 of 61 systems with radial velocity variations fit to orbital solutions, which provide precise measurements of $\period$ and $K$. Fitting the sample's high S/N spectra to templates provides precisely determine $M_1$ values. We use the spectroscopic solutions found in \citet{gianninas14}. Three systems are eclipsing binaries, with known companion masses (J0345$+$1748: $M_2=0.72~\Msun$ \citep{kaplan14}, J0651$+$2844: $M_2=0.50~\Msun$ \citep{brown11b}, and J0751$-$0141: $M_2=0.97~\Msun$ \citep{kilic14}). For these systems, the likelihood reduces to:
\begin{equation}
p(\mf \given \theta) = (1-f_{\rm NS}) \mathcal{N}(M_2^* \given \mu_{\rm WD}, \sigma^2_{\rm WD})
\end{equation}
where $M_2^*$ is the mass of the WD companion in these three systems.

@@

\begin{figure*}[h!]
\begin{center}
\includegraphics[width=0.95\textwidth]{real-data.pdf}
\caption{Results from applying our model to the ELM WDs. The first panel shows the MAP $M_2$ distribution (solid black) and random samples from the posterior (grey lines). The second panel shows posterior samples of $\mu$ and $\sigma$; The MAP Gaussian model has $\mu_{\rm WD}= 0.71~\Msun$ and $\sigma_{\rm WD}= 0.26~\Msun$. The third panel shows posterior samples of $\mu$ and $f_{\rm NS}$. The last panel lists individual LMWDs in the ELM sample (ordered by increasing $\mf$) and shows each of their corresponding NS probabilities. The three systems with all $P_{\rm NS}$ values at 0\% are the three eclipsing systems with known $M_2$'s.}
\label{fig:ELM_post}
\end{center}
\end{figure*}

@@

@@

\begin{thebibliography}{36}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

>>>>

\title{The mass distribution of companions to low-mass white dwarfs}
\author{Jeff J. Andrews\altaffilmark{\colum}, Adrian M. Price-Whelan\altaffilmark{\colum}, Marcel Ag\"ueros\altaffilmark{\colum}}

@@

\begin{abstract}
Measuring the masses of companions to single-line spectroscopic binary stars
is (in general) not possible because of the unknown orbital plane inclination angles. 
Even when the mass of the visible star can be measured, only a lower limit can be placed
on the mass of the unseen companion. However, since these inclination angles should be isotropically
distributed, for a large enough, unbiased sample, the 
companion mass distribution can be deconvolved from the distributions of observables.
In this work, we construct a hierarchical probabilistic model to infer properties of unseen 
companion stars to low mass white dwarfs (LMWD) given orbital period and projected radial velocity measurements of the primary. We use a mixture of 
two Gaussians to model the true companion mass distribution to represent 
populations of white dwarfs (WD) and neutron stars (NS). 
Our model successfully recovers the initial parameters of several mock data sets.
We then apply our model to 55 of the 61 white dwarfs in 
the extremely low mass (ELM) WD survey. We find maximum a posteriori parameters for
the WD companion population of $\mu_{\rm WD} = 0.71~\Msun$ and $\sigma_{\rm WD} = 0.26~\Msun$ 
with a NS fraction consistent with zero, $f_{\rm NS} = 0\%$. However, the marginal posterior 
distribution over NS fraction has a significant tail up to $f_{\rm NS} \approx 16\%$.
We make samples from the posterior distribution publicly available so that future observational
efforts may compute the NS probability for newly discovered low mass WDs.
\end{abstract}

@@

Except in special cases of extreme metallicity \citep{kilic07}, the Galaxy is not old enough to produce LMWDs ($M\lesssim0.45~ \Msun$) through single star evolution. These objects are formed through dynamical interactions with another star, and are therefore expected to be found in binary systems \citep{han98,nelemans00,nelemans01,vdSluys06,woods12}. In their seminal work, \citet{marsh95} discovered companions around five LMWDs in their sample of seven objects, confirming these theoretical expectations. The ELM WD Survey identifies LMWDs in SDSS and elsewhere for spectroscopic follow-up, focusing in particular on extremely low mass WDs ($M\lesssim0.3~ \Msun$) \citep{ELMI}. After their initial investigation found 11 of 12 LMWDs had close binary companions, further spectroscopic follow-up has now brought the sample of ELM WDs to 61 systems, 55 of which have measured radial velocity amplitudes and orbital periods \citep{ELMII, ELMIII, ELMIV, ELMV}.

While the orbital periods and radial velocities indicated the companions were most likely WDs, due to the unmeasured inclination angle, these LMWDs could host NS companions \citep{vLeeuwen07}. However, all radio searches for pulsed emission, as well as searches for blackbody X-ray emission from a NS surface have been unsuccessful \citep{agueros09a,agueros09b,kilic13}. Yet, such systems must exist since LMWDs are observed as companions of millisecond pulsars \citep{vKerkwijk96,callanan98,bassa06,antoniadis12}. Finding even one system would be extremely interesting: the NS binaries with WD companions bright enough to detect spectroscopic radial velocity variations have precisely determined NS masses which provide useful insights into the binary evolution processes that form these systems \citep{benvenuto05,smedley14,jia14}.

%Selecting the most likely candidates for multiwavelength follow-up is crucial. 
For each WD in a single-line spectroscopic binary, spectroscopic observations provide the orbital period ($\period$), primary WD mass ($M_1$), and projected orbital velocity ($K=v \sin i$). Using Kepler's third law, we can write the relation between these quantities and the unknown inclination angle as:
\begin{equation}
	\frac{(M_2 \sin i)^3}{\left(M_1+M_2\right)^2} = \frac{\period}{2\pi G} K^3 = \mf \label{eq:massfunc}
\end{equation}
The righthand side of this equation is the mass function ($\mf$). $M_2$ is minimized for an edge-on orbit, with $i = 90\degree$. Because of the dependence on $i$, the nature of the companion cannot usually be determined based on $\mf$ alone. Figure~\ref{fig:Porb-M1} shows that the population of LMWD with pulsar companions (triangles) occupy the same region in $M_1 - \period$ space as those with WD companions (circles). Therefore, barring rare circumstances (such as an edge-on, eclipsing system), individual LMWDs with NS companions cannot be readily identified.

55 of the 61 systems have precise $M_1$, $\period$, and $K$ measured; we refer to this as the ELM sample. In this work, we develop a probabilistic model to infer parameters of an assumed form for the mass distribution of companions to ELM WDs. Our method is similar to that employed by \citet{ozel12} and \citet{kiziltan13} to describe the mass distribution of NSs in binaries. Instead of spectroscopic fits to component masses, their constraints are based on the measured post-Keplerian parameters in their sample. We focus on several specific astrophysical questions: 1.\ Can the population as a whole be modeled using a simple description of the companion masses? 2.\ How does the companion mass distribution compare to predictions from population synthesis simulations?  3.\ What is the rate of NS-LMWD binaries implied by our model? 4.\ What are the resulting distributions of NS probabilities for individual systems in the ELM sample?

@@

where the product is over each of the $j$ systems.  This marginal likelihood involves integrals over the latent variables $i$ and $M_2$,
\begin{align}
    p(\mf \given \bs{\theta}) &= \int_{-\infty}^\infty dM_2 \int_0^{\pi/2} di  \nonumber \\
      & \qquad {} \times p(\mf \given M_1, M_2, i)~p(M_2 \given \bs{\theta})~p(i).
\end{align}
We neglect observational uncertainties in $\mf$ and $M_1$,\footnote{The fractional uncertainties in these quantities are small, $\sigma_x / x \sim 0.05-0.1$.} and assume the inclination angles are isotropically distributed
\begin{equation}
	p(\mf \given M_1, M_2, i) = \delta \left[\mf - f(M_1, M_2, i) \right]
\end{equation}
where
\begin{equation}
	f(M_1, M_2, i) = \frac{(M_2 \sin i)^3}{(M_1 + M_2)^2}
\end{equation}
and
\begin{equation}
	p(i) = \sin i.
\end{equation}
For now, we do not specify a parametric form for the companion mass distribution, $p(M_2 \given \bs{\theta})$. With the above assumptions, the marginal likelihood integral is
\begin{align}
%    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta}) \int_0^{\pi/2} di \nonumber \\
%    & \qquad {} \times \delta \left[\mf - f(M_1, M_2, i)\right]~\sin i\\
    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2 ~p(M_2 \given \bs{\theta})  \nonumber \\
    & \qquad {} \times \int_0^{\pi/2} di ~\sin i ~ \delta \left[g(M_1,M_2,i) \right]\label{eq:delta}
\end{align}
where
\begin{equation}
	g(M_1,M_2,i) = \mf - \frac{M_2^3}{(M_1+M_2)^2}\sin^3 i.
\end{equation}
The inner integral (over $i$) has the form
\begin{equation}
    \int dx~F(x)~\delta \left[ G(x) \right] = \sum_j \frac{F(x^*_j)}{|G'(x^*_j)|}
\end{equation}
where the sum is over the roots, $x^*_j$, of the function $G(x)$. The root, $i^*$, and derivative of the argument of the delta function in \eqn\ref{eq:delta} are 
\begin{align}
%	\sin i^* &= \alpha\\
%	\alpha &= \frac{(\mf(M_1+M_2)^2)^{1/3}}{M_2}\\
	\sin i^* &= \frac{ \left[\mf(M_1+M_2)^2 \right]^{1/3}}{M_2}\\
%	\frac{\partial g}{\partial i} &= \frac{3M_2^3}{(M_1+M_2)^2}\sin^2 i \cos i\\
%	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \alpha^2 \sqrt{1 - \alpha^2}
	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \sin^2 i^* \sqrt{1 - \sin^2 i^*}
\end{align}
We may rewrite the marginal likelihood as
\begin{align}
	p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\sin i^* \left(\frac{\partial g}{\partial i}\bigg\rvert_{i^*}\right)^{-1}\\
%	&= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\alpha\left[\frac{3M_2^3}{(M_1+M_2)^2} \alpha^2 \sqrt{1 - \alpha^2}\right]^{-1}\\
	&= \int_{M_{2,{\rm min}}}^\infty dM_2~p(M_2 \given \bs{\theta})~h(M_2, \mf, M_1) \label{eq:fullm2}
\end{align}

@@

We must now choose a functional form for the companion mass distribution,  $p(M_2\given \bs{\theta})$. For all experiments -- including that with the real data -- we use a two component Gaussian mixture model to fit the data. In tests below, we generate test data with different mixture component forms. We truncate the distributions using physically motivated bounds so that the white dwarf component is restricted to the range $M_2\in (0.2,\wdupper)~\Msun$ and the neutron star component is restricted to the range $M_2\in (1.3,2)~\Msun$,
\begin{align}
%	h(M_2, \mf, M_1) &= \frac{(M_1+M_2)^{4/3}}{3\mf^{1/3}M_2\sqrt{M_2^2 - \left[ \mf(M_1+M_2)^2 \right]^{2/3}}}\\
%	p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~\left[ (1-f_{\rm NS})p_{\rm WD} + f_{\rm NS}p_{\rm NS} \right]~h(M_2, \mf, M_1)\\
	p(M_2 \given \bs{\theta}) &= \left[ (1-f_{\rm NS})~p_{\rm WD} + f_{\rm NS}~p_{\rm NS} \right] 
\end{align}
where 
\begin{align}
	p_{\rm WD} &= \mathcal{N}(M_2 \given \mu_{\rm WD}, \sigma^2_{\rm WD}); ~0.2 < \frac{M_2}{\Msun} < \wdupper \\
	p_{\rm NS} &= \mathcal{N}(M_2 \given \mu_{\rm NS}, \sigma^2_{\rm NS}); ~1.3 < \frac{M_2}{\Msun} < 2.
\end{align}
and $\mathcal{N}$ is the (truncated, but properly normalized) normal distribution with mean $\mu$ and variance $\sigma^2$ and the distributions are limited to the ranges specified. In this work, we fix the mean and variance of the neutron star distribution to
\begin{align}
	\mu_{\rm NS} &= 1.4~\Msun\\
	\sigma^2_{\rm NS} &= (0.05~\Msun)^2
\end{align}
based on indications that NSs in certain binaries may be somewhat more massive than the canonical NS mass of 1.35 \Msun \citep{kiziltan13,smedley14}.

@@

Our companion mass model parameters are then $\bs{\theta} = (\mu_{\rm WD}, \sigma_{\rm WD}, f_{\rm NS})$. We choose uninformative priors on these parameters: for the mean of the white dwarf component, we use a uniform distribution from $0.2-1.4~\Msun$; we use a (scale-invariant) logarithmic prior for the variance of the white dwarf component over the range $0.05-2~\Msun$; finally, we use a uniform prior over the dimensionless neutron star fraction from $0-1$. The model parameters are summarized in Table~\ref{tbl:parameters}.

@@

We use an ensemble Markov Chain Monte Carlo algorithm \citep{goodman10} to draw samples from the posterior distribution, $p(\mu_{\rm WD}, \sigma_{\rm WD}, f_{\rm NS} \given \bs{m}_f, \bs{M}_1)$. The algorithm is implemented in the \texttt{Python} programming language \citep[{\tt emcee};][]{foremanmackey13} and uses an ensemble of individual ``walkers'' to naturally adapt to the geometry of the parameter-space being explored. We run the walkers for an initial, burn-in period of 500 steps starting from randomly drawn initial conditions (sampled from the priors in Table~\ref{tbl:parameters}). We then re-initialize the walkers from their positions at the end of this run and run again for 1000 steps. We throw out the burn-in samples to eliminate any effects from our choice of initial conditions. The autocorrelation times for each parameter confirm they have converged.

\section{Applying Our Model} \label{sec:tests}
To test the performance of this Gaussian mixture model, we first apply it to three separate sets of mock data (described in detail below), each with 100 ``observed'' systems. These data are generated by randomly drawing the companion mass, $M_2$, from a particular distribution, then using a randomly drawn inclination angle and primary mass, $M_1$, to compute the mass function, $\mf$. We draw primary masses from a uniform distribution, $\mathcal{U}(0.2,0.4)~\Msun$. We apply the same Gaussian mixture model to all test datasets to infer the parameters of the white dwarf mixture component and neutron star fraction.

\begin{figure*}[h!]
\begin{center}
\includegraphics[width=0.95\textwidth]{many-panel.pdf}
\caption{Results from the three tests described in Section~\ref{sec:tests}. The left-most column shows histograms of $M_2$ drawn from each of our test distributions. Overplotted as solid lines are the MAP models. The second column shows samples from the posterior distributions of $\mu_{\rm WD}$ and $\sigma_{\rm WD}$. The dashed lines in the top two panels show the true values that the sample systems were drawn from. Contours designate the 68\% and 95\% contours. Since the third model was not drawn from a Gaussian distribution, there is no true parameter set. The third panel shows posterior samples for $f_{\rm NS}$ and $\sigma_{\rm WD}$, again the dashed lines show the input values. Finally, the rightmost panels list the individual systems (ordered by increasing $\mf$) and show each of their corresponding NS probabilities. }
\label{fig:tests}
\end{center}
\end{figure*}

\subsection{Test 1: Single Gaussian (WD)} \label{sec:exp1}

As a first test, we generate $M_2$ by drawing from a single, truncated Gaussian with parameters given in Table~\ref{tbl:parameters}. This test sample contains no NSs. In the top row of Figure~\ref{fig:ELM_post}, the leftmost panel shows that the maximum a posteriori (MAP) $M_2$ distribution (black line) qualitatively matches the input distribution (grey histogram). The second panel shows samples from the posterior distribution and contours that contain 68\% and 95\% of the samples. The input values lie cleanly within the inner contour in both projections of the posterior shown. The third panel shows that although the posterior $f_{\rm NS}$ distribution is consistent with 0\%, there is a tail up to $\sim$10\%. Nevertheless, our model recovers the input parameters for the white dwarf companion population. The fourth panel lists individual systems, ordered by $\mf$, and shows the distributions of $P_{\rm NS}$ for each. Few systems have $P_{\rm NS}$ values higher than $\sim$5\%.

\subsection{Test 2: Two Gaussians (WD + NS)} \label{sec:exp2}
For our next test, we use the same Gaussian distribution to generate companion masses for the WD component as the previous example, but we add an additional NS component with $f_{\rm NS} = 10\%$ and parameters described in Section~\ref{sec:experiments}. The middle row of Figure~\ref{fig:ELM_post} shows that our model again recovers the input values for $\mu_{\rm WD}$ and $\sigma_{\rm WD}$. Importantly, the third panel shows that it also recovers $f_{\rm NS}$, however the posterior again shows a tail toward higher probabilities. The fourth panel indicates that some systems have near unity $P_{\rm NS}$ values; our model correctly identifies these as true NS, however this accounts for less than half of the systems with NSs in the simulated sample; the other half have low $\mf$'s due to low inclination angles.

\subsection{Test 3: Uniform (WD) + Gaussian (NS)} \label{sec:exp3}
As a final test, we generate companion masses for the WD component by sampling from a uniform distribution over the range $(0.2,1.2)~\Msun$ while keeping $f_{\rm NS}=$10\%. The bottom row of Figure~\ref{fig:ELM_post} shows the results of this test. The posterior distribution in the second panel shows that the systems have difficulty converging on any one solution for $\mu_{\rm WD}$ and $\sigma_{\rm WD}$. Interestingly, the third panel shows that despite having a non-Gaussian input distribution for $M_2$, our model still recovers $f_{\rm NS}$ approximately as accurately as in Test 2.

@@

The ELM WD sample is composed of 55 systems with radial velocity variations fit to orbital solutions, which provide precise measurements of $\period$ and $K$. Fitting the sample's high S/N spectra to templates provides precisely determine $M_1$ values. We use the spectroscopic solutions found in \citet{gianninas14}. Three systems are eclipsing binaries, with known companion masses (NLTT 11748: $M_2=0.72~\Msun$ \citep{kaplan14}, SDSS J065133.3$+$284423.3: $M_2=0.50~\Msun$ \citep{brown11b}, and SDSS J075141.2$-$014120.9: $M_2=0.97~\Msun$ \citep{kilic14}). For these systems, the likelihood reduces to:
\begin{equation}
p(\mf \given \theta) = (1-f_{\rm NS}) \mathcal{N}(M_2^* \given \mu_{\rm WD}, \sigma^2_{\rm WD})
\end{equation}
where $M_2^*$ is the mass of the WD companion in these three systems.

Since the population was selected based on characteristics determined by the primary mass alone, at least with regard to the inclination angle and companion mass, the population is unbiased.

@@

\begin{figure*}[h!]
\begin{center}
\includegraphics[width=0.95\textwidth]{real-data.pdf}
\caption{Results from applying our model to the ELM WDs. The first panel shows the MAP $M_2$ distribution (solid black) and random samples from the posterior (grey lines). The second panel shows posterior samples of $\mu$ and $\sigma$; the MAP Gaussian model has $\mu_{\rm WD}= 0.71~\Msun$ and $\sigma_{\rm WD}= 0.26~\Msun$. The third panel shows posterior samples of $\mu$ and $f_{\rm NS}$. The last panel lists individual LMWDs in the ELM sample (ordered by increasing $\mf$) and shows each of their corresponding NS probabilities. The three systems with all $P_{\rm NS}$ values at 0\% are the three eclipsing systems with measured $M_2$.}
\label{fig:ELM_post}
\end{center}
\end{figure*}

@@

Recently, \citet{hermes14} made high-speed photometric observations of 20 LMWDs in the ELM sample. Based on non-detection of eclipses and modeling of photometric variability, these authors set limits on the inclination angles of these systems. We have not included these observations in our model, but we plan to explore this in the future.

@@

\bibliography{refs}
